\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Risks and Mitigations}\markdownRendererInterblockSeparator
{}Key technical and project risks with practical mitigations.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Hydronic Cooling and Condensation}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: UFH surface and manifolds falling below dew point → condensation, damage.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: hard dew‑point limit controlling mixing valve; diffusion‑tight insulation on cold lines; condensate routing; humidity monitoring per floor; disable cooling if RH persistently > 60\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{System Complexity}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: Dual buffers + FriWa + stove add components and controls.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: safety interlocks in dedicated controllers (not HA); clear operating modes; thorough commissioning; labeled valves/wiring; service documentation.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Noise (Terrace Context)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: Outdoor unit and HRV noise disturbing occupants/neighbors.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: acoustic siting and shielding; resilient mounts; night mode; façade‑friendly HRV placements with baffles.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Thermal Bridges (Loggia, Entry)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: Heat loss, cold surfaces, moisture.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: Prefer enclosure (winter garden/vestibule); otherwise capillary‑active internal insulation and exterior shading; thermography post‑works.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Water Quality and Scaling}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: Scaling and corrosion in plate HX, pumps, valves.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: VDI 2035 treatment; potable‑side filters; periodic checks; bypass/flush ports for service.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Funding/Timing Errors}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: Double‑funding, out‑of‑sequence contracting jeopardizing eligibility.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: EEE involvement; BzA before award; clean cost split; BnD after completion; maintain documentation.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Safety (Stove Integration)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Risk: Overheating without heat dump; low return temp tar formation; insufficient expansion.\markdownRendererUlItemEnd 
\markdownRendererUlItem Mitigations: return‑lift valve (≥60°C), thermal discharge valve to drain, gravity emergency cooling path, expansion vessel sizing; certified components and installation.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight 
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd