\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Funding, Compliance, Documentation}\markdownRendererInterblockSeparator
{}This section outlines a funding approach aligned with current concepts. Always verify current program rules and eligibility with an Energie‑Effizienz‑Experte (EEE) and your bank before committing.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Primary Route — KfW 261 (Wohngebäude – Kredit)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Goal: Effizienzhaus 85 + Erneuerbare‑Energien‑Klasse (≥65- Credit scope: heat pump system, buffers, FriWa, hydraulic periphery, safety, system electricals; potentially wood stove with back‑boiler as part of the system scope.\markdownRendererUlItemEnd 
\markdownRendererUlItem Repayment bonus: typically 10- Important: Costs financed within the EH loan pot cannot be double‑funded elsewhere.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Complementary — Einzelmaßnahmen (BAFA/BEG)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Apply to building envelope (basement/roof/windows/facade), decentralized HRV, measurement/control/automation (MSR), and UFH distribution.\markdownRendererUlItemEnd 
\markdownRendererUlItem Typical rates: 15- Rule: No double‑funding of the same cost position if already in KfW 261 EH pot.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Additional Financing}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Supplementary loans (e.g., KfW programs 358/359) for Einzelmaßnahmen up to common caps per dwelling.\markdownRendererUlItemEnd 
\markdownRendererUlItem PV and battery: 0\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Process and Timing}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Engage an EEE early. Obtain “Bestätigung zum Antrag” (BzA) before awarding contracts, or use suspensive clauses.\markdownRendererUlItemEnd 
\markdownRendererUlItem Clean cost separation: define which positions are under EH‑loan vs. Einzelmaßnahmen to avoid double‑funding.\markdownRendererUlItemEnd 
\markdownRendererUlItem Post‑completion “Bestätigung nach Durchführung” (BnD) to trigger repayment bonus and finalize grants.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Documentation and Compliance}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Design calculations: EN 12831 heat load; ventilation concept per DIN 1946‑6; dew‑point logic; thermal bridge details.\markdownRendererUlItemEnd 
\markdownRendererUlItem Commissioning protocols: blower‑door (pre/post), hydraulic balancing, water quality (VDI 2035), pressure tests, heat meter setup, HP commissioning.\markdownRendererUlItemEnd 
\markdownRendererUlItem Safety: stove per EN 303‑5 with return lift, thermal discharge, emergency cooling; electrical SPDs and backup subpanel documentation.\markdownRendererUlItemEnd 
\markdownRendererUlItem Monitoring: first‑season data logs (temperatures, humidity, heat/power meters) to validate performance.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Notes}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Funding programs evolve. Validate all rates, caps, and eligibility windows at application time.\markdownRendererUlItemEnd 
\markdownRendererUlItem Maintain a single source of truth for cost allocation and supporting documents to simplify audits.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight 
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd