\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Ventilation — Decentralized Single‑Room HRV (DIN 1946‑6)}\markdownRendererInterblockSeparator
{}No central ventilation is feasible; this plan uses 6–8 decentralized HRV units to provide base air exchange with heat recovery, local boosts, and summer bypass.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Design Principles}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Balance and coverage: Provide continuous background ventilation in living/bedrooms; stronger extraction or boost in wet rooms.\markdownRendererUlItemEnd 
\markdownRendererUlItem Acoustic comfort: Low noise at base flow, night mode for bedrooms, acoustic liners/baffles.\markdownRendererUlItemEnd 
\markdownRendererUlItem Simplicity: Short core drill per unit (160–200 mm), slight outward slope for condensate, easy filter access.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Unit Types}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Dual‑fan continuous units (preferred): simultaneous supply and exhaust through a small counter‑flow core; more stable balance.\markdownRendererUlItemEnd 
\markdownRendererUlItem Alternating push‑pull units (paired): ceramic core stores heat; install in pairs to approximate balance.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Placement and Air Paths}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Wet rooms (bath, WC, kitchen): units with boost 40–60 m³/h; grease filters where needed.\markdownRendererUlItemEnd 
\markdownRendererUlItem Living/bedrooms: base 15–30 m³/h per room; night mode in bedrooms.\markdownRendererUlItemEnd 
\markdownRendererUlItem Door undercuts/transfer grilles to enable crossflow; avoid short‑circuiting supply to immediate exhaust.\markdownRendererUlItemEnd 
\markdownRendererUlItem Exterior intakes with weather hoods and insect screens; consider façade acoustics.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Controls}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Local control: manual boost switches (bath fans), humidity/CO₂‑based automatic boosts, night modes.\markdownRendererUlItemEnd 
\markdownRendererUlItem Central overview (optional): dry contacts or Modbus/IP bridge to integrate with Home Assistant for visibility and logging (not safety‑critical).\markdownRendererUlItemEnd 
\markdownRendererUlItem Summer bypass: enable to prevent unwanted heat recovery during cooling season; querventilation possible.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Installation and Commissioning}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Core drills with slight outward slope for condensate; ensure airtight sleeve sealing.\markdownRendererUlItemEnd 
\markdownRendererUlItem Configure design flow rates per DIN 1946‑6 ventilation concept; measure and record flows.\markdownRendererUlItemEnd 
\markdownRendererUlItem Filters: set maintenance intervals; keep spare sets; include filter‑service notifications.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Interaction with Hydronic Cooling}\markdownRendererInterblockSeparator
{}\markdownRendererUlBegin
\markdownRendererUlItem HRV aids humidity control but does not dehumidify aggressively; monitor RH per floor.\markdownRendererUlItemEnd 
\markdownRendererUlEnd \markdownRendererInterblockSeparator
{}- If RH > 60
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd