\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Controls, Sensors, Monitoring}\markdownRendererInterblockSeparator
{}Safety controls must be hard‑wired; Home Assistant (HA) is used for orchestration and monitoring only. Dew‑point protection is mandatory for UFH cooling.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Safety/Primary Controls (Hard‑wired)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Heat pump controller: compressor protections, defrost logic, supply/return limits.\markdownRendererUlItemEnd 
\markdownRendererUlItem Hydronic controller: 3‑way mixing valve control for UFH, pump control, buffer temperature limits, DHW priority.\markdownRendererUlItemEnd 
\markdownRendererUlItem Dew‑point cutout: direct input limiting UFH supply temperature to ≥(dew point + 2 K); independent of HA.\markdownRendererUlItemEnd 
\markdownRendererUlItem Stove safety: return‑lift valve (≥60°C), thermal discharge safety valve, gravity emergency cooling path; pressure relief; expansion vessel sizing.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Sensors and Instrumentation}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Buffers: top/mid/bottom temperatures on DHW and heating/cooling buffers.\markdownRendererUlItemEnd 
\markdownRendererUlItem UFH: supply and return temperatures; optional manifold surface temperature.\markdownRendererUlItemEnd 
\markdownRendererUlItem Ambient: at least one temperature/RH sensor per floor for dew point.\markdownRendererUlItemEnd 
\markdownRendererUlItem Energy metering: heat meters for AWHP→buffer and FriWa primary (optional), electric sub‑meter for HP.\markdownRendererUlItemEnd 
\markdownRendererUlItem Flow/pressure: as required by FriWa and UFH balancing; include purge points and drains.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Orchestration (Home Assistant or similar)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem PV‑aware schedules: DHW buffer charge at midday; optional winter buffer preheat during PV surplus.\markdownRendererUlItemEnd 
\markdownRendererUlItem Seasonal modes: winter (heating), summer (cooling + DHW only), shoulder (adaptive minimal setpoints).\markdownRendererUlItemEnd 
\markdownRendererUlItem Notifications: wood stove opportunity prompts; filter service (HRV/FriWa); abnormal RH/temperature alerts; fault relays from HP.\markdownRendererUlItemEnd 
\markdownRendererUlItem Data logging: buffer temps, UFH VL/RL, indoor RH/T, HP power, heat meter data; derive coefficient of performance (COP) proxy where meters available.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Example Dew‑Point Logic}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Inputs: indoor T, indoor RH → dew point; measured UFH supply temperature.\markdownRendererUlItemEnd 
\markdownRendererUlItem Limit: UFH supply target = max(heating curve demand, dew point + 2 K) with absolute min typically 19–21°C.\markdownRendererUlItemEnd 
\markdownRendererUlItem Fault/lockout: if supply < (dew point + 2 K) or manifold surface moisture detected, close mixing valve/stop pump until conditions safe.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Electrical and Resilience}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Critical‑loads subpanel: HP enable, circulators, hydronic controller, HA host, network gear, fridge, selected lights/outlets.\markdownRendererUlItemEnd 
\markdownRendererUlItem Surge protection: SPD coordination for PV, battery, HP, control electronics.\markdownRendererUlItemEnd 
\markdownRendererUlItem Night mode: reduce acoustic output of outdoor unit and HRV where possible.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight 
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd