\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Commissioning and Acceptance}\markdownRendererInterblockSeparator
{}This checklist defines the procedures and evidence required to safely commission and accept the system. It is structured to match funding documentation needs.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Pre‑Commissioning}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Mechanical:\markdownRendererUlItemEnd 
\markdownRendererUlItem Pipework flushed, pressure tested; leak‑free.\markdownRendererUlItemEnd 
\markdownRendererUlItem Expansion vessels sized and pre‑charged for total volume (buffers + circuits).\markdownRendererUlItemEnd 
\markdownRendererUlItem Air/dirt separators installed at strategic points; vents accessible.\markdownRendererUlItemEnd 
\markdownRendererUlItem Insulation complete: hot lines insulated; cold lines diffusion‑tight with condensate routes.\markdownRendererUlItemEnd 
\markdownRendererUlItem Electrical:\markdownRendererUlItemEnd 
\markdownRendererUlItem Critical‑loads subpanel wired; circuits labeled; RCD/MCB verified.\markdownRendererUlItemEnd 
\markdownRendererUlItem Surge protection devices (SPDs) installed and coordinated with PV/battery.\markdownRendererUlItemEnd 
\markdownRendererUlItem Control wiring (sensors, valves, pumps) labeled; emergency stops documented.\markdownRendererUlItemEnd 
\markdownRendererUlItem Water quality:\markdownRendererUlItemEnd 
\markdownRendererUlItem System water prepared per VDI 2035; conductivity/hardness logged.\markdownRendererUlItemEnd 
\markdownRendererUlItem Safety (stove):\markdownRendererUlItemEnd 
\markdownRendererUlItem Return‑lift valve operation verified; thermal discharge safety connected to drain; gravity emergency cooling path tested; chimney approvals.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Controls and Sensors Validation}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Buffers: confirm top/mid/bottom sensor readings; directions of flow validated.\markdownRendererUlItemEnd 
\markdownRendererUlItem UFH: verify supply/return sensors; 3‑way valve orientation; pump rotation.\markdownRendererUlItemEnd 
\markdownRendererUlItem Dew‑point logic: inject high RH in test zone or simulate via controller; confirm UFH supply limit and lockout behavior.\markdownRendererUlItemEnd 
\markdownRendererUlItem HP controller: weather curve loaded; limits set (min/max VL, anti‑short‑cycle).\markdownRendererUlItemEnd 
\markdownRendererUlItem FriWa: outlet temperature control tuned; stable performance over 2–16 L/min draws; anti‑scald verified.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Functional Tests}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Heating mode (winter simulation):\markdownRendererUlItemEnd 
\markdownRendererUlItem Long steady compressor operation; buffer stratification observed; target VL achieved at test outdoor setpoint.\markdownRendererUlItemEnd 
\markdownRendererUlItem DHW midday priority: buffer top reaches 55–60°C; resume space heating post‑charge.\markdownRendererUlItemEnd 
\markdownRendererUlItem Cooling mode (summer simulation):\markdownRendererUlItemEnd 
\markdownRendererUlItem Night buffer chill to ~16–18°C; daytime UFH VL maintained at dew point + 2 K; no condensation on manifolds/lines.\markdownRendererUlItemEnd 
\markdownRendererUlItem Stove integration:\markdownRendererUlItemEnd 
\markdownRendererUlItem Charge buffer top; verify AWHP priority reduction on high top‑of‑buffer temperature.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Hydraulic Balancing}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem UFH circuits: measure and record volume flows per loop; adjust to design; document setpoints.\markdownRendererUlItemEnd 
\markdownRendererUlItem Heat meters: verify installation direction and pulse outputs if logging.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Ventilation (Decentralized HRV)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Design flows set per room; boost functions verified; summer bypass configured.\markdownRendererUlItemEnd 
\markdownRendererUlItem Filters installed; maintenance calendar established; acoustic checks.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Acceptance Documentation}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Blower‑door test results (pre + post) with achieved n50 ≤ 1.5 h⁻¹ (target).\markdownRendererUlItemEnd 
\markdownRendererUlItem Hydronic balancing protocol (flows per loop), HP and FriWa commissioning reports.\markdownRendererUlItemEnd 
\markdownRendererUlItem VDI 2035 water quality report; pressure test certificates; electrical test reports (RCD/insulation/SPD).\markdownRendererUlItemEnd 
\markdownRendererUlItem Schematics: final hydraulic diagram, control I/O map, sensor list.\markdownRendererUlItemEnd 
\markdownRendererUlItem Safety attestations: stove compliance (EN 303‑5), chimney sweep approvals, emergency procedures.\markdownRendererUlItemEnd 
\markdownRendererUlItem Monitoring plan: which data is logged and retention period (first season minimum).\markdownRendererUlItemEnd 
\markdownRendererUlEndTight 
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd