\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Bill of Materials (Specification Classes)}\markdownRendererInterblockSeparator
{}Brand‑agnostic list with size ranges. Final selections should follow detailed calcs and installer standards.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Heat Generation and Storage}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Reversible AWHP (R290), 6–10 kW class, monobloc or split; low noise kit; night mode.\markdownRendererUlItemEnd 
\markdownRendererUlItem Optional Plate Heat Exchanger (AWHP↔House), sized for full capacity with low approach ΔT; glycol on AWHP side.\markdownRendererUlItemEnd 
\markdownRendererUlItem Heating/Cooling Buffer: 800–1000 L, stratified, 100–150 mm insulation, 3× thermowells.\markdownRendererUlItemEnd 
\markdownRendererUlItem DHW Heating Buffer: 200–300 L, high insulation, 3× thermowells.\markdownRendererUlItemEnd 
\markdownRendererUlItem FriWa Module: 25–35 kW plate HX, variable‑speed primary pump, flow sensor, electronic outlet control, service valves, potable filter/strainer.\markdownRendererUlItemEnd 
\markdownRendererUlItem Wood Stove w/Back‑Boiler (optional): rated output matched to buffer; return‑lift valve (≥60°C), thermal discharge valve, emergency cooling path components; chimney parts as required.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Hydronic Periphery}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Circulation Pumps: ECM pumps for AWHP loop(s), buffer charging, UFH circuits, FriWa primary.\markdownRendererUlItemEnd 
\markdownRendererUlItem Valves: 3‑way mixing valve (UFH), motorized zone valves for seasonal decoupling, check valves, balancing valves, drain/fill valves.\markdownRendererUlItemEnd 
\markdownRendererUlItem Separators: air and dirt separators at key locations; magnetic dirt separator if needed.\markdownRendererUlItemEnd 
\markdownRendererUlItem Expansion Vessels: sized for total water volume; service valves and gauges.\markdownRendererUlItemEnd 
\markdownRendererUlItem Safety: PRVs (typically 3 bar), manometers, automatic air vents; condensate traps for cold lines.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Distribution}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem UFH Manifolds and Loops: oxygen‑barrier PEX/MLCP, manifold cabinets, flow meters, actuators if zoned.\markdownRendererUlItemEnd 
\markdownRendererUlItem Pipe Insulation: hot lines to code; cold lines diffusion‑tight; manifold/valve box insulation where possible.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Ventilation (Decentralized HRV)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem 6–8 single‑room HRV units (dual‑fan continuous or push‑pull pairs), wall sleeves, exterior hoods, acoustic liners, filters.\markdownRendererUlItemEnd 
\markdownRendererUlItem Control accessories: boost switches, humidity/CO₂ sensors (where supported), integration gateway (optional).\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Controls, Sensors, Electrical}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Hydronic Controller: mixing valve + pump control with dew‑point input and lockout.\markdownRendererUlItemEnd 
\markdownRendererUlItem Sensors: buffer temps (top/mid/bottom), UFH VL/RL, ambient T/RH per floor, optional manifold surface probe.\markdownRendererUlItemEnd 
\markdownRendererUlItem Energy Meters: heat meter(s) on AWHP and FriWa primary (optional), sub‑meter for HP electrical.\markdownRendererUlItemEnd 
\markdownRendererUlItem Electrical: critical‑loads subpanel, automatic transfer switch (ATS) for backup, surge protective devices (SPDs), labeling, wiring accessories.\markdownRendererUlItemEnd 
\markdownRendererUlItem Home Assistant Host: small, reliable compute (e.g., SBC or mini‑PC), network connectivity, UPS (optional).\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Water Treatment and Service}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem VDI 2035 treatment unit/chemistry, test kit (conductivity/hardness), fill/drain assemblies.\markdownRendererUlItemEnd 
\markdownRendererUlItem Filters/Strainers: potable filter for FriWa; strainers where needed on primary; spare filter sets.\markdownRendererUlItemEnd 
\markdownRendererUlItem Serviceability: isolation valves, drain points, thermowells/test points, access panels.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight 
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd