\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Project Phases and Recommended Order}\markdownRendererInterblockSeparator
{}A practical sequence minimizes rework, aligns with funding steps, and protects commissioning quality.\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Phases}\markdownRendererInterblockSeparator
{}1) Pre‑checks and Surveys - Hazard screening (1972 build): asbestos in adhesives/fibre‑cement, PCB, old mineral wool. - Roof statics for PV; outdoor unit siting and acoustics; local permits (loggia/vestibule).\markdownRendererInterblockSeparator
{}2) Concept and Calculations - EN 12831 heat and (if needed) cooling loads; DIN 1946‑6 ventilation concept. - Thermal‑bridge detailing at loggia and entry; dew‑point strategy for UFH cooling. - Define cost allocation to EH loan vs. Einzelmaßnahmen.\markdownRendererInterblockSeparator
{}3) Funding Applications - EEE issues BzA (confirmation before application) and supports bank discussion. - Submit KfW 261 EH loan; plan complementary Einzelmaßnahmen (BAFA/BEG) and any supplementary loans.\markdownRendererInterblockSeparator
{}4) Envelope and Airtightness - Basement ceiling, roof/attic, windows; optional facade. - Stair partition; loggia/vestibule if approved. - Blower‑door test (intermediate if feasible) for QA.\markdownRendererInterblockSeparator
{}5) Technical Systems - UFH installation and distribution; hydraulic balancing readiness. - AWHP, buffers (DHW + heating/cooling), FriWa, pumps/valves, sensors, electrical subpanel. - Decentralized HRV units with commissioning of design flows.\markdownRendererInterblockSeparator
{}6) PV and Battery - PV array and battery install; integrate critical‑loads subpanel; SPDs.\markdownRendererInterblockSeparator
{}7) Commissioning and Tuning - Flush, pressure test, VDI 2035; parameterization; dew‑point test; stove safety tests. - Hydraulic balancing protocol; data logging setup.\markdownRendererInterblockSeparator
{}8) Post‑Completion - EEE issues BnD (confirmation after completion); repayment bonus/grants processed. - First‑season monitoring; optimize setpoints; perform thermography if needed.
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd