\markdownRendererDocumentBegin
\markdownRendererSectionBegin
\markdownRendererHeadingOne{Envelope, Airtightness, Thermal Bridges}\markdownRendererInterblockSeparator
{}This section defines measures for insulation, airtightness, and mitigation of key thermal bridges (loggia, open entrance/vestibule, open staircase to cellar).\markdownRendererInterblockSeparator
{}\markdownRendererSectionBegin
\markdownRendererHeadingTwo{Targets (EH85‑aligned)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Basement ceiling U ≤ 0.25 W/(m²·K)\markdownRendererUlItemEnd 
\markdownRendererUlItem Roof/attic U ≤ 0.14 W/(m²·K)\markdownRendererUlItemEnd 
\markdownRendererUlItem Windows Uw ≤ 0.90 W/(m²·K) with warm edge spacers\markdownRendererUlItemEnd 
\markdownRendererUlItem Airtightness n50 ≤ 1.5 h⁻¹ (blower‑door before/after)\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Measures}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Basement ceiling insulation (must): rigid boards or blown‑in solutions; ensure continuous coverage around beams and services; seal penetrations.\markdownRendererUlItemEnd 
\markdownRendererUlItem Roof/attic insulation (must): on‑roof insulation or upper ceiling; maintain continuous airtight layer; detail penetrations and edges.\markdownRendererUlItemEnd 
\markdownRendererUlItem Windows (must): triple glazing, airtight installation with tapes and compressible seals; verify reveal insulation.\markdownRendererUlItemEnd 
\markdownRendererUlItem Facade (optional, recommended): external WDVS where feasible; if not, capillary‑active internal insulation (e.g., CaSi/wood fibre) at critical interior surfaces; model details to avoid vapor traps.\markdownRendererUlItemEnd 
\markdownRendererUlItem Airtightness package (must): professional sealing, tape schedules, and blower‑door test pre/post.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Staircase to Cellar (Open Stair)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Install airtight glazed partition/door at basement or ground‑floor landing; maintain natural light; perimeter seals.\markdownRendererUlItemEnd 
\markdownRendererUlItem Combine with basement ceiling insulation to cut stack‑effect losses.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Loggia (Interior Balcony) Options}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Winter garden conversion: high‑performance triple glazing, thermally broken frames, insulated parapets/soffits, exterior shading; best comfort and thermal‑bridge fix; requires structural and permit checks.\markdownRendererUlItemEnd 
\markdownRendererUlItem Targeted thermal‑bridge remediation: capillary‑active internal insulation at junctions; add exterior insulation if feasible; less impact but lower cost.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Open Entrance / Covered Entry (Below Living Space)}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Glazed vestibule (Windfang): insulated framing and external door; major infiltration/comfort improvement; may require permit.\markdownRendererUlItemEnd 
\markdownRendererUlItem Underside insulation: rigid PIR/EPS below ceiling; seal side‑wall junctions; economical but smaller impact.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight \markdownRendererInterblockSeparator
{}
\markdownRendererSectionEnd \markdownRendererSectionBegin
\markdownRendererHeadingTwo{Verification}\markdownRendererInterblockSeparator
{}\markdownRendererUlBeginTight
\markdownRendererUlItem Thermography after completion to identify residual thermal bridges.\markdownRendererUlItemEnd 
\markdownRendererUlItem Blower‑door test (n50 ≤ 1.5 h⁻¹) with documented leakage sealing actions.\markdownRendererUlItemEnd 
\markdownRendererUlEndTight 
\markdownRendererSectionEnd 
\markdownRendererSectionEnd \markdownRendererDocumentEnd